Acceleration of global search through dual Lipschitz constant estimates

### 2019-03-07 2019/NUMTA_Barkalov/Abstract/Barkalov.tex
Second English version

\title{A new way to speedup in global optimization problems}
\author{Roman Strongin, Konstantin Barkalov, Semen Bevzuk}

@@

\textbf{Keywords.} Global optimization; multiextremal problems; Lipschitz constant.

@@

The paper considers global optimization problems with a black-box objective function satisfying a Lipschitz condition. 
An important issue in this class of problems is reliable estimates of a priori unknown Lipschitz constant based on the results of the search trials. 
In fundamental work \cite{Book} various approaches have been proposed to take into account both global and local properties of the objective function. 
%В частности, показали свою перспективность  алгоритмы с использованием локальных оценок константы Липшица 
In particular, algorithms using local estimates of the Lipschitz constant have shown their promise \cite{Sergeyev10, Sergeyev16}.

This paper proposes a new way to account for the local properties of the objective function.
The proposed method is based on the use in the information-statistical global search algorithm \cite{Book} lower and upper estimates of the Lipschitz constant simultaneously. 
The rules of the new algorithm provide automatic switching between these two estimates, depending on the type of the objective function and on the search stage.

Proposed algorithm guarantees convergence to global minimizer. The results of solving several series of multiextremal test problems confirm both the global convergence of the new algorithm and its acceleration compared to the original one.

>>>>

\title{Acceleration of global search through dual Lipschitz constant estimates}
\author{Roman Strongin, Konstantin Barkalov, Semen Bevzuk}

@@

\textbf{Keywords.} Global optimization; multiextremal problems; Lipschitz constant estimates.

@@

The paper considers global optimization problems with a black-box objective function satisfying the Lipschitz condition. 
Efficient algorithms for this class of problems require reliable estimates of the Lipschitz constant to be introduced. 
In fundamental work \cite{Book} various approaches have been proposed to take into account both global and local properties of the objective function. 
%В частности, показали свою перспективность  алгоритмы с использованием локальных оценок константы Липшица 
In particular, algorithms using local estimates of the Lipschitz constant have shown their potential \cite{Sergeyev10, Sergeyev16}.

The new approach presented in this paper is based on simultaneous use of two estimates: one is substantially larger then the other. 
The larger estimate ensures global convergence and the smaller one reduces the total number of trials needed to find the global optimizer.
Results of numerical experiments on the random sample of multidimensional functions demonstrate the efficiency of the suggested approach.
